\thetitle

### 2008-05-29 paper.tex
Fixed bibliography problems.

It is already common practice in network modeling to think of flow behaviors as vectors.
The \caps{IP} protocol number, source and destination \caps{IP} addresses, and \caps{UDP/TCP} source and port numbers that define a packet flow are commonly referred to as a 5-tuple---i.e., a vector in five dimensions.
Other aspects of a flow's behavior can be represented vectorially as well. We call such a mapping of flows into a vector space a \emph{linear representation} if it satisfies the following simple linearity criterion:
\begin{quote}
The sum of the representations of two flows must be another vector that represents the aggregate behavior of the two flows.
\end{quote}
We show that flows can be represented satisfying this criterion. This immediately allows us to perform useful computations on traffic patterns through simple vector addition and scaling.
Finally, we argue that the natural generalization of this approach is to represent an entire network's behavior as a large matrix, with each row representing a single flow.
The matrix representation gives a representation of network behavior that preserves the non-uniformity and non-independence of flow behaviors, but still is highly amenable to analysis using theoretical and numerical linear algebra.
Moreover, we describe how to generate traffic from arbitrary traffic matrices, and demonstrate experimentally that the specific matrix representation we propose preserves the performance characteristics of real traffic.

@@

The history of networking research contains many examples of simplistic models that have proven not only to be inaccurate, but also to drastically skew important characteristics of network behavior. Paxson and Floyd showed that the Poisson packet arrival model, which had been standard for studying wide-area Internet traffic, failed to capture the burstiness and self-similarity of real traffic~\cite{Paxson95}. The equally common but simplistic Random Way-Point~(\caps{RWP}) mobility model was found to exhibit ``density waves'' and gradual slow-down of average node speed~\cite{Royer01,Yoon03:speed-decay}. In the worst cases, overly simplistic models can switch the relative performance of protocols, thereby invalidating the conclusions drawn from performance comparisons using those models.

The interaction of wireless users and application behavior with the lower layers of the networking stack is characterized by where, when, how much, and to whom data is transmitted. The joint pattern of traffic generation and mobility through time and space completely determines the effect of wireless usage on the lower levels of the network. This is due to the data-agnostic nature of the protocol stack: by design, \caps{IP} networks treat all data in the same manner.%~\cite{Clark88}.
\footnote{This is violated by some quality of service (QoS) schemes. However, we can simply add QoS metadata---such as traffic classes or urgency flags---to our models of user behavior and the rest of our arguments remain valid. The network is still disinterested in the exact content of the data being transported; only the QoS metadata is relevant.} The credibility of conclusions derived from simulation or experimental deployment depends crucially on our confidence that the models used to generate traffic in experiments are sufficiently realistic.

@@

%\pagebreak
% Hernandez06:dissertation: ``complete methodology for reproducing the traffic observed on a network link in a closed-loop manner, and proposed a number of metrics for studying the realism of the generated traffic.''
\bibliography{IEEE,references}

>>>>

It is already common practice in network modeling to think of flow behaviors as vectors.
The \caps{IP} protocol number, source and destination \caps{IP} addresses, and \caps{UDP/TCP} source and port numbers that define a packet flow are commonly referred to as a 5-tuple---i.e., a vector in five dimensions.
Other aspects of a flow's behavior can be represented vectorially as well. We call such a mapping of flows into a vector space a \emph{linear representation} if it satisfies the following simple linearity criterion:
\begin{quote}
The sum of the representations of two flows is a vector that represents the aggregate behavior of the two flows.
\end{quote}
We show that flows can be represented satisfying this criterion. This immediately allows us to perform useful computations on traffic patterns through simple vector addition and scaling.
Finally, we argue that the natural generalization of this approach is to represent an entire network's behavior as a large matrix, with each row representing a single flow.
The matrix representation gives a representation of network behavior that preserves the non-uniformity and non-independence of flow behaviors, but still is highly amenable to analysis using theoretical and numerical linear algebra.
Moreover, we describe how to generate traffic from arbitrary traffic matrices, and demonstrate experimentally that the specific matrix representation we propose preserves the performance characteristics of real traffic.

@@

%The history of networking research contains many examples of simplistic models that have proven not only to be inaccurate, but also to drastically skew important characteristics of network behavior. Paxson and Floyd showed that the Poisson packet arrival model, which had been standard for studying wide-area Internet traffic, failed to capture the burstiness and self-similarity of real traffic~\cite{Paxson95}. The equally common but simplistic Random Way-Point~(\caps{RWP}) mobility model was found to exhibit ``density waves'' and gradual slow-down of average node speed~\cite{Royer01,Yoon03:speed-decay}. In the worst cases, overly simplistic models can switch the relative performance of protocols, thereby invalidating the conclusions drawn from performance comparisons using those models.

\newfootnote{\QoSnote}{This is violated by some quality of service (QoS) schemes. However, we can simply add QoS metadata---such as traffic classes or urgency flags---to our models of user behavior and the rest of our arguments remain valid. The network is still disinterested in the exact content of the data being transported; only the QoS metadata is relevant.}

The interaction of wireless users and application behavior with the lower layers of the networking stack is characterized by where, when, how much, and to whom data is transmitted. The joint pattern of traffic generation and mobility through time and space completely determines the effect of wireless usage on the lower levels of the network. This is due to the data-agnostic nature of the protocol stack: by design, \caps{IP} networks treat all data in the same manner~\cite{Clark88}.\QoSnote The credibility of conclusions derived from simulation or experimental deployment depends crucially on our confidence that the models used to generate traffic in experiments are sufficiently realistic.

@@

%\pagebreak
% Hernandez06:dissertation: ``complete methodology for reproducing the traffic observed on a network link in a closed-loop manner, and proposed a number of metrics for studying the realism of the generated traffic.''
\bibliographystyle{plain}
\bibliography{references}
